---
title: "Tüseb"
author: "beyza"
date: "2022-12-08"
output: html_document
---

load libraries
```{r}
library(argparse)
library(findpython)
library(sp)
library(SeuratObject)
library(Seurat)
library(hdf5r)
library(devtools)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(patchwork)
library(devtools)
library(AUCell)
library(Seurat)
```



```{r}
getwd()
setwd("/Users/beyzaaygun/Desktop/Bioinformatic/TÜSEB/GSE122960")
```


SeuratObject oluşturmak için
```{r}

GSM3489182_obj <- Read10X_h5(filename = "GSM3489182_Donor_01_filtered_gene_bc_matrices_h5.h5",
                             use.names = TRUE,
                             unique.features = TRUE)



Seurat_GSM3489182 <- CreateSeuratObject(counts = GSM3489182_obj)


```





Seurat
1.MTReads
```{r}
Seurat_GSM3489182[["percent.mt"]] <- PercentageFeatureSet(Seurat_GSM3489182, pattern = "^MT-")
VlnPlot(Seurat_GSM3489182, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```


2. filtering
```{r}
Seurat_GSM3489182 <- subset(Seurat_GSM3489182, subset =nFeature_RNA > 200 & nFeature_RNA <4500 & percent.mt < 17)
```

3. Normalize data
```{r}
#pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)

Seurat_GSM3489182 <- NormalizeData(Seurat_GSM3489182)
```


4. Identify highly variable features
```{r}
Seurat_GSM3489182 <- FindVariableFeatures(Seurat_GSM3489182, selection.method = "vst", nfeatures = 3863)
```


5. Scaling
```{r}
all.genes <- rownames(Seurat_GSM3489182)
Seurat_GSM3489182 <- ScaleData(Seurat_GSM3489182, features = all.genes)
```


6.Perform linear dimensionality reduction
```{r}
Seurat_GSM3489182 <- RunPCA(Seurat_GSM3489182, features = VariableFeatures(object = Seurat_GSM3489182))
```


EV ilgili genleri expression seviyesini görmek için
```{r}
features <- c("CD9", "CD63", "CD81")
RidgePlot(Seurat_GSM3489182, features = features, ncol = 1)
VlnPlot(Seurat_GSM3489182, features = features)
```
tsne
```{r}
Seurat_GSM3489182 <- RunTSNE(
  Seurat_GSM3489182,
  reduction = "pca",
  cells = NULL,
  dims = 1:5,
  features = NULL,
  seed.use = 1,
  tsne.method = "Rtsne",
  dim.embed = 2,
  distance.matrix = NULL,
  reduction.name = "tsne",
  reduction.key = "tSNE_"
)

DimPlot(Seurat_GSM3489182, reduction = "tsne")
```




install AUCell
```{r}
devtools::install_github("aertslab/AUCell", build_vignettes=F)
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

BiocManager::install("AUCell")
```



For GeneSetrs
```{r}
genes <- unique(c("CD9", "CD63", "CD81", "PDCD6IP", "TSG101","PDCD6IP","HSP60", "HSP70", "HSP90", "CD82", "CD37","VTA1", "VPS4", "VPS32", "ARF6", "ANXA1"))
geneSets <- GeneSet(genes, setName="geneSet2")
exprMatrix <- Seurat_GSM3489182
```


1. Score gene signatures
RUN YOK
```{r}

counts <- GetAssayData(object = exprMatrix, slot = "counts")

```



1.1. Build gene-expression rankings for each cell
```{r}
cells_rankings <- AUCell_buildRankings(counts)
cells_rankings
```



1.2. Calculate enrichment for the gene signatures (AUC)
```{r}
cells_AUC <- AUCell_calcAUC(geneSets, cells_rankings)
```



2. Determine the cells with the given gene signatures or active gene sets
2. neden boş????
```{r}
cells_assignment <- AUCell_exploreThresholds(cells_AUC, plotHist=TRUE, assign=TRUE)

cells_assignment$Oligodendrocyte_Cahoy$aucThr$thresholds
```


```{r}
library("ggplot2")
geneSetName <- rownames(cells_AUC)[grep("geneSet2", rownames(cells_AUC))]
AUCell_plotHist(cells_AUC[geneSetName,], aucThr=0.35)
abline(v=0.39)

```


```{r}
quartz()
newSelectedCells <- names(which(getAUC(cells_AUC)["geneSet2",]>0.35))

length(newSelectedCells)
```


```{r}
library("Seurat")
exprMatrix$is_ec <- ifelse(colnames(exprMatrix) %in% newSelectedCells, "EV", "non_EV")

DimPlot(reduction = "tsne",object=exprMatrix, group.by = "is_ec", label = TRUE)
```


Derinlik3
```{r}
selectedThresholds <- getThresholdSelected(cells_assignment)
par(mfrow=c(2,3)) # Splits the plot into two rows and three columns
for(geneSetName in names(selectedThresholds))
{
  nBreaks <- 5 # Number of levels in the color palettes
  # Color palette for the cells that do not pass the threshold
  colorPal_Neg <- grDevices::colorRampPalette(c("black","blue", "skyblue"))(nBreaks)
  colorPal_Neg._num <- as.numeric(colorPal_Neg)
  # Color palette for the cells that pass the threshold
  colorPal_Pos <- grDevices::colorRampPalette(c("pink", "magenta", "red"))(nBreaks)
  
  # Split cells according to their AUC value for the gene set
  passThreshold <- cells_AUC[geneSetName,] >  selectedThresholds[geneSetName]
  if(sum(passThreshold) >0 )
  {
     aucSplit <- split(cells_AUC[geneSetName,], passThreshold)

    # Assign cell color
    cellColor <- c(setNames(colorPal_Neg[cut(aucSplit[[1]], breaks=nBreaks)], names(aucSplit[[1]])), 
                  setNames(colorPal_Pos[cut(aucSplit[[2]], breaks=nBreaks)], names(aucSplit[[2]])))
    
    # Plot
    plot(exprMatrix$Y, main=geneSetName,
         sub="Pink/red cells pass the threshold",
         col=cellColor[rownames(exprMatrix$Y)], pch=16) 
  }
}
```












