---
title: "Tüseb"
author: "beyza"
date: "2022-12-08"
output: html_document
---

load libraries
```{r}
library(argparse)
library(findpython)
library(sp)
library(SeuratObject)
library(Seurat)
library(hdf5r)
library(devtools)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(patchwork)
library(devtools)
library(AUCell)
library(Seurat)
```



```{r}
getwd()
setwd("/Users/beyzaaygun/Desktop/Bioinformatic/TÜSEB/GSE122960")
```


SeuratObject oluşturmak için
```{r}

GSM3489182_obj <- Read10X_h5(filename = "GSM3489182_Donor_01_filtered_gene_bc_matrices_h5.h5",
                             use.names = TRUE,
                             unique.features = TRUE)



Seurat_GSM3489182 <- CreateSeuratObject(counts = GSM3489182_obj)


```





Seurat
1.MTReads
```{r}
Seurat_GSM3489182[["percent.mt"]] <- PercentageFeatureSet(Seurat_GSM3489182, pattern = "^MT-")
VlnPlot(Seurat_GSM3489182, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```


2. filtering
```{r}
Seurat_GSM3489182 <- subset(Seurat_GSM3489182, subset =nFeature_RNA > 200 & nFeature_RNA <4500 & percent.mt < 17)
```

3. Normalize data
```{r}
#pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)

Seurat_GSM3489182 <- NormalizeData(Seurat_GSM3489182)
```


4. Identify highly variable features
```{r}
Seurat_GSM3489182 <- FindVariableFeatures(Seurat_GSM3489182, selection.method = "vst", nfeatures = 3863)
```


5. Scaling
```{r}
all.genes <- rownames(Seurat_GSM3489182)
Seurat_GSM3489182 <- ScaleData(Seurat_GSM3489182, features = all.genes)
```


6.Perform linear dimensionality reduction
```{r}
Seurat_GSM3489182 <- RunPCA(Seurat_GSM3489182, features = VariableFeatures(object = Seurat_GSM3489182))
```


EV ilgili genleri expression seviyesini görmek için
```{r}
features <- c("CD9", "CD63", "CD81")
RidgePlot(Seurat_GSM3489182, features = features, ncol = 1)
VlnPlot(Seurat_GSM3489182, features = features)
```


install AUCell
```{r}
devtools::install_github("aertslab/AUCell", build_vignettes=F)
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

BiocManager::install("AUCell")
```



For GeneSetrs
```{r}
genes <- c("CD9", "CD63", "CD81")
geneSets <- list(geneSet1=genes)
exprMatrix <- Seurat_GSM3489182
```


1. Score gene signatures
RUN YOK
```{r}

counts <- GetAssayData(object = exprMatrix, slot = "counts")

```



1.1. Build gene-expression rankings for each cell
```{r}
cells_rankings <- AUCell_buildRankings(counts)
cells_rankings
```



1.2. Calculate enrichment for the gene signatures (AUC)
```{r}
cells_AUC <- AUCell_calcAUC(geneSets, cells_rankings)
```



2. Determine the cells with the given gene signatures or active gene sets
2. neden boş????
```{r}
cells_assignment <- AUCell_exploreThresholds(cells_AUC, plotHist=TRUE, assign=TRUE)

cells_assignment$Oligodendrocyte_Cahoy$aucThr$thresholds
```



Explore cells/clusters based on the signature score
```{r}
load(paste(file.path(system.file('examples', package='AUCell')), "cellsTsne.RData", sep="/"))
cellsTsne <- cellsTsne$Y
plot(cellsTsne, pch=16, cex=.3)
```














