---
title: "Tüseb"
author: "beyza"
date: "2022-12-08"
output: html_document
---


load libraries
```{r}
library(argparse)
library(findpython)
library(sp)
library(SeuratObject)
library(Seurat)
library(hdf5r)
library(devtools)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(patchwork)
library(devtools)
library(AUCell)
library(Seurat)
library(GSEABase)
library(Rtsne)


```



```{r}
getwd()
setwd("~/Desktop/Bioinformatic/TÜSEB/GSE122960")
```





SeuratObject oluşturmak için
```{r}

GSM3489182_obj <- Read10X_h5(filename = "/Users/beyzaaygun/Desktop/Bioinformatic/TÜSEB/GSE122960/GSM3489182_Donor_01_filtered_gene_bc_matrices_h5.h5", 
                             use.names = TRUE,
                             unique.features = TRUE)

Seurat_GSM3489182 <- CreateSeuratObject(counts = GSM3489182_obj)

```


Seurat
1.MTReads
```{r}
Seurat_GSM3489182[["percent.mt"]] <- PercentageFeatureSet(Seurat_GSM3489182, pattern = "^MT-")

VlnPlot(Seurat_GSM3489182, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)


```

2. filtering
```{r}
Seurat_GSM3489182 <- subset(Seurat_GSM3489182, subset =nFeature_RNA > 200 & nFeature_RNA <4500 & percent.mt < 17)
```

3. Normalize data
```{r}
#pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)
Seurat_GSM3489182 <- NormalizeData(Seurat_GSM3489182)
```


4. Identify highly variable features
```{r}
Seurat_GSM3489182 <- FindVariableFeatures(Seurat_GSM3489182, selection.method = "vst", nfeatures = 3863)
```


5. Scaling
```{r}
all.genes <- rownames(Seurat_GSM3489182)
Seurat_GSM3489182 <- ScaleData(Seurat_GSM3489182, features = all.genes)
```


6.Perform linear dimensionality reduction
```{r}
Seurat_GSM3489182 <- RunPCA(Seurat_GSM3489182, features = VariableFeatures(object = Seurat_GSM3489182))
Seurat_GSM3489182
```


7. umap
```{r}

Seurat_GSM3489182 <- FindNeighbors(Seurat_GSM3489182, dims = 1:10)
Seurat_GSM3489182 <- FindClusters(Seurat_GSM3489182, resolution = 0.5)
Seurat_GSM3489182 <- RunUMAP(Seurat_GSM3489182, dims = 1:15)
DimPlot(Seurat_GSM3489182, reduction = "umap")


```



EV ilgili genleri expression seviyesini görmek için
```{r}
features <- c("CD9", "CD63", "CD81", "PDCD6IP", "TSG101", "TSPAN6", "TSPAN8")
RidgePlot(Seurat_GSM3489182, features = features, ncol = 1)
VlnPlot(Seurat_GSM3489182, features = features)

```



For GeneSetrs
"CD9", "CD63", "CD81", "PDCD6IP", "TSG101", "ARF6", "ANXA1", "CD82", "CD37", "VTA1", "CD80", "CD86", "SIRT6", "SDCBP", "TRIM5", "CLTC", "VPS4A", "VPS4B", "HSPA4", "HSPA1A", "HSPA4", "HSPA8", "HSPA1B" ,"HSPA5", "HSPA14", "HSPA2", "HSPA9", "HSPD1", "HSP90AA1", "HSP90AB1", "PLD1", "PLD2", "CD40", "CHMP4B", "CHMP4A", "CHMP4C", "FLOT1", "FLOT2", "ANXA2", "ANXA6", "ANXA4", "ANXA7", "CFL1", "CFL2", "PFN1", "PFN2"

"CD9", "CD63", "CD81", "PDCD6IP", "TSG101", "GAPDH", "SDCBP", "SDCBP2", "UBC", "CLTC", "CLTA", "CHMP4B", "CHMP4C", "VPS4A", "VPS4B", "MAPK1", "MAPK3", "ARF6"
```{r}
genes <- unique(c("CD9", "CD63", "CD81", "PDCD6IP", "TSG101", "ARF6", "ANXA1", "CD82", "CD37", "VTA1", "CD80", "CD86", "SIRT6", "SDCBP", "TRIM5", "CLTC", "VPS4A", "VPS4B", "HSPA4", "HSPA1A", "HSPA4", "HSPA8", "HSPA1B" ,"HSPA5", "HSPA14", "HSPA2", "HSPA9", "HSPD1", "HSP90AA1", "HSP90AB1", "PLD1", "PLD2", "CD40", "CHMP4B", "CHMP4A", "CHMP4C", "FLOT1", "FLOT2", "ANXA2", "ANXA6", "ANXA4", "ANXA7", "CFL1", "CFL2", "PFN1", "PFN2"))
geneSets <- GeneSet(genes, setName="geneSet1")
exprMatrix <- Seurat_GSM3489182
exprMatrix_A <- Seurat_GSM3489182@assays$RNA@counts

cells_AUC <- AUCell_run(exprMatrix_A, geneSets)
```





```{r}
counts <- GetAssayData(object = exprMatrix, slot = "counts")
cells_rankings <- AUCell_buildRankings(counts)
cells_rankings
```




```{r}
cells_AUC <- AUCell_calcAUC(geneSets, cells_rankings)
```




2. Determine the cells with the given gene signatures or active gene sets

```{r}
cells_assignment <- AUCell_exploreThresholds(cells_AUC, plotHist=TRUE, assign=TRUE)
cells_assignment$Oligodendrocyte_Cahoy$aucThr$thresholds

```


tekli yapınca hata???????????

```{r}
library("ggplot2")
geneSetName <- rownames(cells_AUC)[grep("geneSet1", rownames(cells_AUC))]
AUCell_plotHist(cells_AUC[geneSetName,], aucThr=0.23)
abline(v=0.23)


```


```{r}
library("ggplot2")
geneSetName <- rownames(cells_AUC)[grep("geneSet1", rownames(cells_AUC))]
AUCell_plotHist(cells_AUC[geneSetName,], aucThr=0.23)
abline(v=0.23)


```



```{r}
quartz()
newSelectedCells <- names(which(getAUC(cells_AUC)["geneSet1",]>0.23))

length(newSelectedCells)
```


```{r}


exprMatrix$is_ec <- ifelse(colnames(exprMatrix) %in% newSelectedCells, "EV", "non_EV")

DimPlot(reduction = "umap",object=exprMatrix, group.by = "is_ec", label = TRUE)












